---
title: "Assignment 2 INFO-659"
author: "Rishabh Sharma"
output:
  html_document:
    df_print: paged
---


# A. Data Preparation

## A.1. Data Loading and Initial Transformation
```{r}
df <- read.csv("UCI_Credit_Card (1).csv")
View(df)
```

```{r}
head(df, 5)
```
```{r}
df$default.payment.next.month <- factor(df$default.payment.next.month,levels=c(0,1), labels=c("No","Yes"))
```

Other than these transformation, We need to do additional transofrmations on the sex, marriage, PAY_0, PAY_1, PAY_2, PAY,3, PAY_4, PAY_5 and PAY_6

## A.4. Transforming Nominal Variables
```{r}
df$SEX <- factor(df$SEX,levels=c(1,2), labels=c("male", "female"))
df$MARRIAGE <- factor(df$MARRIAGE,levels=c(1,2,3), labels=c("married", "single", "other")) 
df$EDUCATION <- factor(df$EDUCATION,levels=c(1,2,3,4,5,6), labels=c("graduate school", "university", "high school","others","unknown","unknown"))

df$PAY_0 <- factor(df$PAY_0,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))

df$PAY_2 <- factor(df$PAY_2,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))

df$PAY_3 <- factor(df$PAY_3,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))
df$PAY_4 <- factor(df$PAY_4,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))
df$PAY_5 <- factor(df$PAY_5,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))
df$PAY_6 <- factor(df$PAY_6,levels=c(-2,-1,0,1,2,3,4,5,6,7,8), labels=c("pay duly", "pay duly", "pay duly",
                                                                        "delay 1 mo","delay 2 mo","delay 3 mo",
                                                                        "delay 4 mo","delay 5 mo","delay 6 mo",
                                                                        "delay 7 mo","delay 8 mo"))


```

After these changes lets check what our dataset looks like

```{r}
head(df,10)
```

## A.2. Demographic Variables

```{r}
library(ggplot2)
ggplot(df, aes(x=EDUCATION, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack") +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("darkolivegreen4", "red"))
```


From this we can see that most of the data that we have is from population who have an university level education. With most of the people not having a default status next month. 

#### Plotting sex against frequency for default
```{r}
ggplot(df, aes(x=SEX, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack") +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("white", "brown"))
```

This graph shows us that we have more data for the female populus. Also,maximum female population has no Default payment for the next month.

```{r}
ggplot(df, aes(x=MARRIAGE, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack") +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("orange", "blue"))
```

This graph shows us that most of the population is single which has contributed to the data. Also, Married and singles have almost similar number of defaulters for the next month, Staying close to around 2,600

Now plotting around the age,

```{r}
ggplot(df, aes(x=AGE, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("darkolivegreen4", "red"))
```

Here we can see a skewed binomial distribution. We can probably change it to a bell curve by a log transformation. 

Here we can see that bell curve for The defaulters and total population distributed by the age, have somewhat same skewed nature. Thus we can draw the assumption, That most of the defaulters came from the age of 22 to 45 and then declined steadily.

Also there must be some correlation between the Yeses for defaulters in next month and age
```{r}
ggplot(df, aes(x=log(AGE), fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("darkolivegreen4", "red"))
```

As we did not have a varied data in the age section, Log transformation didn't perform quite well. 


## A.3. Payment Status Variables

Now to plot for PAY_0, PAY_1, PAY_2, PAY,3, PAY_4, PAY_5 and PAY_6

```{r}
ggplot(df, aes(x=PAY_0, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("orange", "blue"))
```

Here most of our data is concentrated for people who duly pay(largest chunk of data) with some data in for people who are late by one, two and three months. Although we can still see that people who have duly paid the bills have yes as defaulters for the next month.
```{r}
ggplot(df, aes(x=PAY_2, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("darkolivegreen4", "red"))
```

Here most of the populus has paid their installments duly, but there is some data for people who have delayed their payment by two months.

```{r}
ggplot(df, aes(x=PAY_3, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("green", "white"))
```

Just like PAY_2, Here most of the populus has paid their installments duly, but there is some data for people who have delayed their payment by two months.


```{r}
ggplot(df, aes(x=PAY_4, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("yellow", "brown"))
```

Just like PAY_2 and PAY_3, Here most of the populus has paid their installments duly, but there is some data for people who have delayed their payment by two months.

```{r}
ggplot(df, aes(x=PAY_5, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("yellow", "blue"))
```



This graphical representation seems to showing that the most people delayed by 2 months

```{r}
ggplot(df, aes(x=PAY_6, fill=default.payment.next.month, color=default.payment.next.month)) + 
  geom_bar(width = 0.5, position="stack")  +
  scale_color_manual(values=c("black","black")) +
  scale_fill_manual(values=c("grey", "red"))
```

```{r}
typeof(df$PAY_0)
```


### hypthesis formulation

From the above data we can say that we can predict the defaulters in next month can be predicted using classification algorithms using attributes such as age(because of the correlation sobserved in the graph), sex, Education, marriage and the Payment features. 

This can be a good business solution if one can predict which of the users are going to default in the upcoming month so that the company can target such customers and take some preventive measures.
 
Running the models.
## A.5. Selection of Training Data

```{r}
train <- df[sample(nrow(df), 5000), ]
```
## A.6. Selection of Testing Data

```{r}
test <- df[c(1,214,23),]
```

# B. Data Classification

## B.1. Naive Bayes using Demographic Variables
```{r}
library(e1071)
nbDem <- naiveBayes(default.payment.next.month ~ SEX + EDUCATION + MARRIAGE, train)
nbDem
```
For 0, our probability is 0.7782 and 1 is o.2218 which makes sense, since in our intial analysis we came to know that there are more non defaulters than the defaulters. 


Running our trained model on testing data set

```{r}
predict(nbDem, test[0,])
```

```{r}
predict(nbDem, test[2,])
```

## B.2. Naive Bayes using Payment Status
```{r}
library(e1071)
nbDem <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3 + PAY_4 +PAY_5 +PAY_6, train)
nbDem
```
```{r}
predict(nbDem, test[1,])
```

```{r}
predict(nbDem, test[2,])
```

## B.3. Smoothed Naive Bayes using Payment Status

```{r}
nbPay <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3, train, laplace=1.5)
predict(nbPay, test[1,])
```
Changing the laplace value to 1 and 2
```{r}
nbPay <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3, train, laplace=1)
predict(nbPay, test[3,])
```

```{r}
nbPay <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3, train, laplace=2.0)
predict(nbPay, test[1,])
```
# C. Classification with Decision Tree

## C.1. Basic Decision Tree


```{r}
library("rpart")
library("rpart.plot")
dtPay <- rpart(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6,
            method="class",
            data=train, parms=list(split='information'), 
            minsplit=20, cp=0.04)
```

```{r}
predict(dtPay, test[1,])
```

```{r}
predict(dtPay, test[2,])

```
```{r}
predict(dtPay, test[3,])

```


```{r}
rpart.plot(dtPay, type=4, extra=1)
```

## C.2. Decision Tree with a Different Complexity Parameter


```{r}
dtPay <- rpart(default.payment.next.month ~ PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6,
            method="class",
            data=train, parms=list(split='information'), 
            minsplit=20, cp=0.01)
rpart.plot(dtPay, type=4, extra=1)
```


```{r}
predict(dtPay, test[1,])
```

```{r}
predict(dtPay, test[2,])

```
```{r}
predict(dtPay, test[3,])

```

# D. Conclusion
Since, The accuracy was not calculated, It would be very hard to share the true results. but one can clearly say that both the models worked pretty well. 

Naive Bayse showed us that the number of elements for which we will be getting No would be some what close to 78% whereas, Yeses would be close to 22%. Which is the conclusion one can draw from the data exploration done in the part B of this notebook. 

Decision Tree although gives us a better approach on how we can know how the decision was made by the model. Apart from this decision tree told us about the probablity of yes and no for each test case, Something naive bayes failed to do. 

# ASSIGNMENT 3

## A.1 
```{r}
ggplot(dfx, aes(as.factor(default.payment.next.month), LIMIT_BAL)) + 
    geom_point() + 
    labs(y = "limit balance", x = "default");
```


```{r}
typeof(df$LIMIT_BAL)
```



From this we can say that most of the defaulters stayed in the range of 0 to 750000. Where as customers with no defaults have a range from 0 to somewhat close to 800000. Since, it is a comparision from a categorical variable with a continuous data no such evident differences can be found. 


For comparing PAY_0,2,3,4,5, and 6. We did additional EDA to draw relationships between the pay and the next month default status. (from line 125 to 180) under section A3 assignment 2.

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT1, fill = default.payment.next.month)) + 
    geom_bar(stat='identity') 
```
```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT2, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT3, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```


```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT4, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT5, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```



```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = BILL_AMT6, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
typeof(df$BILL_AMT1)
```


From the relationship of various bill amounts for various months along with the bill amount status, we can see that. They follow a very similar trend when it comes to defaulters and non defaulters. We can see that defaulters for all the bill amounts are lower than the non defaulters. also they lie in the range of 0 to 3.0e+08. 

We can clearly say that the defaulters for next months had a lower bill amount than the people who did not have a default in the next month. This can be one of the reasons they became defaulters in the next month.

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT1, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```
```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT2, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT3 , fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT4, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT5, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = PAY_AMT6, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```

```{r}
typeof(df$PAY_AMT1)
```

A similar trend can be seen for the Pay amount. The non defaulters had their pay amounts as high as 1.7e+08 where as it was very low for the defaulters. Thus meaning that people who can be defaulters next month can be classified by using the trends of their lower bill amount and pay amount.


## A. 2

From analyzing some of the data in A1 we can clearly see that, PAY_AMT and BILL_AMT can be succesfully used to draw a conclusion for the defaulters in the next month. It can be seen that for the defaulters bill amounts and the pay amounts for the defaulters were significantly lower than the customers who did not default their credit card bills. 

#### 1. making a new attribute using all the pay amounts and all the bill amounts. So that we can find what is the total pay that the customer is liable to pay in the next month. 

```{r}
df$TOTAL_DUE <- (abs(df$BILL_AMT1)) - (abs(df$PAY_AMT1)) + (abs(df$BILL_AMT2)) - (abs(df$PAY_AMT2)) + (abs(df$BILL_AMT3)) - (abs(df$PAY_AMT3)) + (abs(df$BILL_AMT4)) - (abs(df$PAY_AMT4)) + (abs(df$BILL_AMT5)) - (abs(df$PAY_AMT5)) + (abs(df$BILL_AMT6)) - (abs(df$PAY_AMT6))
```


```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = TOTAL_DUE, fill = default.payment.next.month)) + 
    geom_bar(stat='identity')
```



#### 2. making a new attribute as ratio of limit balance along with the sum of all the bill amounts

```{r}
df$LIMIT_TO_BILL_RATIO <- abs(df$LIMIT_BAL) / ((abs(df$BILL_AMT1)) + (abs(df$BILL_AMT2)) + (abs(df$BILL_AMT3)) + (abs(df$BILL_AMT4)) + (abs(df$BILL_AMT5)) + (abs(df$BILL_AMT6)))
```


```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = LIMIT_TO_BILL_RATIO, fill = default.payment.next.month)) + 
    geom_point(stat='identity')
```
#### 3.  making a new attribute as ratio of limit balance along with the sum of all the pay amounts

```{r}
df$LIMIT_TO_PAY_RATIO <- abs(df$LIMIT_BAL) / ((abs(df$PAY_AMT1)) + (abs(df$PAY_AMT2)) + (abs(df$PAY_AMT3)) + (abs(df$PAY_AMT4)) + (abs(df$PAY_AMT5)) + (abs(df$PAY_AMT6)))
```

```{r}
ggplot(data = df, aes(x = default.payment.next.month, y = LIMIT_TO_PAY_RATIO, fill = default.payment.next.month)) + 
    geom_point(stat='identity')
```
On a superficial level, Both the graphs, i.e. limit to pay ratio and limit to bill ratio with respect to the default status graphs look the same. But they can be helpful to notice various relationships to identify if the person is going to default the credit next month. 


## B. Data Preperation and Cleansing

#### B.1 Since the data is already loaded and ready to run operations on, With all the necessary transformations done in the assignment 3. 

#### B.2 Create a filtered dataset with only non-negative amounts

```{r}
dfnn <- subset(df, BILL_AMT1>=0 & BILL_AMT2>=0 &BILL_AMT3>=0 & BILL_AMT4>=0 &BILL_AMT5>=0 & BILL_AMT6>=0 & PAY_AMT1>=0 & PAY_AMT2>=0 & PAY_AMT3>=0 & PAY_AMT4>=0 & PAY_AMT5>=0 & PAY_AMT6>=0)
nrow(dfnn)
```

```{r}
View(dfnn)
```

## C. Data Transformation and Classification/Modeling 

Using naive bayes as the model to form a calssification prediction. 


segregating 90% of the data for the training and 10% of the data for testing randomly.
```{r}
library(caTools)

set.seed(1)

sample <- sample.split(dfnn, SplitRatio = 0.9)
train  <- subset(dfnn, sample == TRUE)
test   <- subset(dfnn, sample == FALSE)

train <- train[!is.infinite(rowSums(train)),]
test  <- test[!is.infinite(rowSums(test)),]


```

```{r}
nbDem <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6 + TOTAL_DUE , train)
nbDem
```

```{r}
prediction = predict(nbDem, newdata=test)
```


```{r}
table(prediction, test$default.payment.next.month)
```

```{r}
install.packages("caret", dependencies = c("Depends", "Suggests"))
```
```{r}
library(caret)

confusion <- confusionMatrix(prediction ,as.factor(test$default.payment.next.month))

```

```{r}
precision <- confusion$byClass['Pos Pred Value']    
recall <- confusion$byClass['Sensitivity']

```
 
```{r}
confusion
```

```{r}
f_measure <- 2 * ((precision * recall) / (precision + recall))
```

```{r}
f_measure
```

```{r}
precision
```

```{r}
recall
```

## C.2 

```{r}
nbDem <- naiveBayes(default.payment.next.month ~ PAY_0 + PAY_2 + PAY_3 + PAY_4 + PAY_5 + PAY_6+ LIMIT_TO_BILL_RATIO + LIMIT_TO_PAY_RATIO, train)

nbDem
```

```{r}
View(test)

prediction <- predict(nbDem, newdata=test)
prediction
confusion <- confusionMatrix(prediction ,as.factor(test$default.payment.next.month))
```

```{r}
precision <- confusion$byClass['Pos Pred Value']    
recall <- confusion$byClass['Sensitivity']

confusion
```

```{r}
f_measure <- 2 * ((precision * recall) / (precision + recall))
 

f_measure
```

```{r}
precision
```
```{r}
recall
```

## C3. 
```{r}
hist(log2(dfnn$TOTAL_DUE))
```
```{r}
train$TOTAL_DUE_LOG <- (log(train$TOTAL_DUE))
test$TOTAL_DUE_LOG <- (log(test$TOTAL_DUE))

train <- train[!is.infinite(rowSums(train)),]
test  <- test[!is.infinite(rowSums(test)),]

nbDem <- naiveBayes(default.payment.next.month ~ BILL_AMT1 + BILL_AMT2 + BILL_AMT3 + BILL_AMT4 + BILL_AMT5 + BILL_AMT6 + LIMIT_TO_BILL_RATIO + LIMIT_TO_PAY_RATIO + TOTAL_DUE_LOG, train)

nbDem
```

```{r}
prediction <- predict(nbDem, newdata=test)

confusion <- confusionMatrix(prediction ,as.factor(test$default.payment.next.month))
```

```{r}
precision <- confusion$byClass['Pos Pred Value']    
recall <- confusion$byClass['Sensitivity']

confusion
```

```{r}
f_measure <- 2 * ((precision * recall) / (precision + recall))

f_measure
```

```{r}
precision
```

```{r}
recall
```

## C4. 

```{r}
nbDem <- naiveBayes(default.payment.next.month ~ LIMIT_TO_BILL_RATIO + LIMIT_TO_PAY_RATIO + TOTAL_DUE_LOG, train)

nbDem
```

```{r}
prediction <- predict(nbDem, newdata=test)

confusion <- confusionMatrix(prediction ,as.factor(test$default.payment.next.month))
```

```{r}
precision <- confusion$byClass['Pos Pred Value']    
recall <- confusion$byClass['Sensitivity']

confusion
```

```{r}
f_measure <- 2 * ((precision * recall) / (precision + recall))

f_measure
```


```{r}
precision
```

```{r}
recall
```


## D Evaluation and result

With the accuracy, precision, recall, f1 score, and kappa calculated, This part can be answered. All the models along with various parameters used along with the transformed parameters gave satisfactory results. Proving our hypothesis discussed in the assignment 2(conclusion).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

